Chr2 hotspot preliminary analyses

### 2019-06-24 analysis/paper/Chr2-prelim.Rmd
updated Chr2 prelim analysis code

>>>>

## Filter to remove local traits that are outside hotspot

In initial analyses for each hotspot, I included too many local traits. Namely, I considered all local traits that are within 7 Mb of the hotspot midpoint. I wish to limit this to traits that are within 3 Mb of the hotspot midpoint.

```{r define_foo}
foo <- mylist %>%
  bind_rows() %>%
  rename(pleiotropy_lod = lod) %>%
  left_join(tib, by = c("local_gene_id", "nonlocal_gene_id")) %>%
  select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff, lod_diff_proportion, lod_without_med, lod_with_med, filename, nonlocal_gene_id, local_gene_id) 
```

```{r add_local_peak_positions}
foo %>%
  left_join(local_annot2, by = c("local_gene_id" = "annot.id")) %>%
  rename(local_peak_pos = pos, local_peak_lod = lod) %>%
  select(- marker.id, - chrom, - strand, - biotype, - local_tx) %>%
  filter(local_peak_pos < hot_upper, local_peak_pos > hot_lower)

```{r}
foo %>%
  left_join(local_annot2, by = c("local_gene_id" = "annot.id")) %>%
  rename(local_peak_pos = pos, local_peak_lod = lod) %>%
  select(- marker.id, - chrom, - strand, - biotype, - local_tx) %>%
  filter(local_peak_pos < hot_upper, local_peak_pos > hot_lower) %>%
  select(middle) %>% unlist() %>% range()

## Convert tibble to matrix for clustering of pleiotropy test statistics

```{r, hclust_pleio_stats}
bar <- foo %>%
  select(nonlocal_gene_id, local_gene_id, pleiotropy_lod) %>%
hotspots::tibble_to_matrix() %>%
  dist() %>%
  hclust() 
```

```{r load_colors}
library(RColorBrewer) 
hmcol <- colorRampPalette(brewer.pal(9, "GnBu"))(100)
```

```{r heat_pleio_stats}
foo %>%
  select(nonlocal_gene_id, local_gene_id, pleiotropy_lod, lod_diff) %>%
hotspots::tibble_to_matrix() %>%
  t() %>%
  gplots::heatmap.2(cexRow = 0.1, cexCol = 0.1, col = hmcol, trace = "none")
```

```{r heat_lod_diff}
foo %>%
  select(nonlocal_gene_id, local_gene_id, lod_diff) %>%
hotspots::tibble_to_matrix() %>%
  t() %>%
  gplots::heatmap.2(cexRow = 0.1, cexCol = 0.1, col = hmcol, trace = "none")
```

```{r heat_lod_diff0}
foo %>%
  select(nonlocal_gene_id, local_gene_id, lod_diff) %>%
  mutate(lod_diff = lod_diff * (lod_diff > 0)) %>%
hotspots::tibble_to_matrix() %>%
  t() %>%
  gplots::heatmap.2(cexRow = 0.1, cexCol = 0.1, col = hmcol, trace = "none")
```

```{r, define_hld, eval = FALSE}
hld <- foo %>%
  select(nonlocal_gene_id, local_gene_id, lod_diff)
plotly::plot_ly(as.data.frame(hld), z = as.data.frame(hld)$lod_diff, type = "heatmap")
```

## Compare dendograms

Should we create a lod difference dendogram based on the mediations with a single local trait? Or with the full matrix of mediation analyses?

```{r load_dendextend}
suppressPackageStartupMessages(library(dendextend))
```

```{r define_hc_objects}
hc_lod_diff <- foo %>%
  select(nonlocal_gene_id, local_gene_id, lod_diff) %>%
  hotspots::tibble_to_matrix() %>%
  dist() %>%
  hclust() 
hc_pleio <- foo %>%
  select(nonlocal_gene_id, local_gene_id, pleiotropy_lod) %>%
  hotspots::tibble_to_matrix() %>%
  dist() %>%
  hclust() 
```

```{r load_dendextend}
library(dendextend)
```

```{r dend_dist}
dend_diff(as.dendrogram(hc_pleio), as.dendrogram(hc_lod_diff))
```

## 2 x 2 tables

We binarize the continuum of mediation LOD difference values and that of pleiotropy test statistics.

```{r 2x2_table}
mediation_threshold <- 1.5
mediation_threshold_proportion <- 0.5
pleiotropy_threshold <- 4
table(med = foo$lod_diff > mediation_threshold, pleio = foo$pleiotropy_lod <= pleiotropy_threshold)
```

```{r 2x2_table_proportion}
table(med = foo$lod_diff_proportion > mediation_threshold_proportion, pleio = foo$pleiotropy_lod <= pleiotropy_threshold)
```

```{r 2x2_table_proportion}
table(med = foo$lod_diff_proportion > 0.4, pleio = foo$pleiotropy_lod <= pleiotropy_threshold)
```

```{r counts_lod_diff}
foo %>%
  filter(lod_diff > mediation_threshold, pleiotropy_lod <= pleiotropy_threshold) %>%
  group_by(local_symbol) %>%
  count() %>%
  arrange(desc(n))

```{r counts_lod_diff_prop_first}
foo %>%
  filter(lod_diff_proportion > mediation_threshold_proportion, pleiotropy_lod <= pleiotropy_threshold) %>%
  group_by(local_symbol) %>%
  count() %>%
  arrange(desc(n))

```{r counts_both}
foo %>%
  filter(lod_diff > mediation_threshold, lod_diff_proportion > mediation_threshold_proportion, pleiotropy_lod <= pleiotropy_threshold) %>%
  group_by(local_symbol) %>%
  count() %>%
  arrange(desc(n))
```

```{r counts_either}
foo %>%
  filter(lod_diff_proportion > mediation_threshold_proportion | lod_diff > mediation_threshold, pleiotropy_lod <= pleiotropy_threshold) %>%
  group_by(local_symbol) %>%
  count() %>%
  arrange(desc(n))

```{r counts_lod_diff_prop}
foo %>%
  filter(lod_diff_proportion > 0.2, pleiotropy_lod <= pleiotropy_threshold) %>%
  group_by(local_symbol) %>%
  count() %>%
  arrange(desc(n))

```{r counts_more}
foo %>%
  filter(lod_diff_proportion > 0.2, pleiotropy_lod <= pleiotropy_threshold) %>%
  group_by(nonlocal_symbol) %>%
  count() %>%
  arrange(desc(n))
foo %>% 
  filter(lod_diff_proportion > 0.2, pleiotropy_lod <= pleiotropy_threshold) %>%
  select(local_symbol, nonlocal_symbol, pleiotropy_lod, lod_diff_proportion, lod_diff)
  arrange(desc(nonlocal_symbol)) 
```
